HTW EXAM sims

### 2023-11-09 Simulations/htw_exam.qmd
exam fitting

```{r}
# load and view data
pacman::p_load(tidyverse,lme4,future,furrr,patchwork,here)
purrr::walk(here(c("Functions/Display_Functions.R", "Functions/alm_core.R")),source)
select <- dplyr::select; mutate <- dplyr::mutate

d <- readRDS(here("data/e1_08-21-23.rds"))
levels(d$condit)
dtest <- d %>% filter(expMode %in% c("test-Nf","test-train-nf")) %>% group_by(id,lowBound) %>% 
  mutate(nBand=n(),band=bandInt,id=factor(id)) %>% group_by(id) %>% mutate(nd=n_distinct(lowBound))
dtest <- dtest %>% group_by(id,lowBound) %>% filter(nBand>=5 & nd==6)
dtest <- dtest %>% group_by(id) %>% filter(!id %in% unique(dtest$id[dtest$nBand<5]))

ds <- d %>% filter(expMode2 %in% c("Train","Test")) |> 
  filter(!id %in% unique(dtest$id[dtest$nBand<5])) |>   
  group_by(id,condit,expMode2) |> 
  # create tr as trial number within id, condit, expMode2, 
  mutate(input=bandInt,x=bandInt, y=vx, tr= row_number()) |>
  select(id,condit,expMode2,tr,x,y) 
  #select(id,condit,expMode2,trial,gt.train,vb,band,bandInt,input,vx,dist,vxb)

# dst <- ds %>% filter(expMode2=="Train")
# dst <- dst %>%
#   group_by(id, vb) %>%
#   mutate(trial_band = row_number())

# create dsAvg, averaging over id, within tr
dsAvg <- ds |> group_by(condit,expMode2,tr, x) |> 
  summarise(y=mean(y),.groups="keep")

@@

@@

inNodes <- 7; outNodes = 14; 
input.layer <- seq(50,1600,length.out=inNodes*1)  
output.layer <- seq(50,1600,length.out=outNodes*1)

weights = matrix(.0000,nrow=length(output.layer),ncol=length(input.layer))

adjust_layer <- function(input.layer, k){
  if(k == 1) return(input.layer)
  new.layer <- c()
  for(i in 1:(length(input.layer) - 1)){
    new.layer <- c(new.layer, seq(input.layer[i], input.layer[i+1], length.out = k + 1))
  }
  new.layer <- unique(new.layer)
  return(new.layer)
}

default.layer <- c(100,350,600,800,1000,1200)
k = 2
input.layer <- adjust_layer(default.layer, k)
output.layer <- adjust_layer(default.layer, k)

@@

optimize_params <- function(dsAvg, c_values, lr_values, input.layer, output.layer){
    best_error <- Inf
    best_params <- list(c = NA, lr = NA)

    # Split the data into training and testing sets
    train_data <- dsAvg[dsAvg$expMode2 == "Train", ]

    test_data <- dsAvg[dsAvg$expMode2 == "Test", ] |> group_by(x) |> summarise(y=mean(y))

    # Grid search over c and lr values
    for (c in c_values) {
        for (lr in lr_values) {
            # Train ALM with current parameters
            sim_results <- alm.sim(train_data, c, lr, input.layer, output.layer)
            weight.mat <- sim_results$wm
            trainVec <- c(0,unique(train_data$x))
            # Test on the testing data
            predictions <- sapply(test_data$x, function(input) {
                exam.response(input, c,trainVec, input.layer, output.layer, weight.mat)
            })
            (test_data <- cbind(test_data, pred=predictions))
            # Compute mean squared error
            mse <- mean((predictions - test_data$y)^2)

        # Update best parameters if current error is lower
            if (!is.na(mse) && mse < best_error) {
                best_error <- mse
                best_params$c <- c
                best_params$lr <- lr }
        }
    }
    print(best_params)
    print(best_error)

    return(best_params)
  }

# Example usage
c_values <- seq(0.000001, 0.01, length.out=40)
lr_values <- seq(0.01, 1.5, length.out=20)
best_paramsV <- optimize_params(dsAvg |> filter(condit=="Varied"), c_values, lr_values, input.layer, output.layer)
c=best_paramsV$c; lr=best_paramsV$lr

best_paramsC <- optimize_params(dsAvg |> filter(condit=="Constant"), c_values, lr_values, input.layer, output.layer)
c=best_paramsC$c; lr=best_paramsC$lr

@@

## EXAM - trial wise
```{r}

optimize_params <- function(dsAvg, c_values, lr_values, input.layer, output.layer){
    best_error <- Inf
    best_params <- list(c = NA, lr = NA)

    # Split the data into training and testing sets
    train_data <- dsAvg[dsAvg$expMode2 == "Train", ]
    test_data <- dsAvg[dsAvg$expMode2 == "Test", ] |> group_by(x) |> summarise(y = mean(y))

    # Determine the second half of the training data
    train_data_half <- train_data[(round(nrow(train_data)/2) + 1):nrow(train_data), ]

    # Grid search over c and lr values
    for (c in c_values) {
        for (lr in lr_values) {
            # Train ALM with current parameters
            sim_results <- alm.sim(train_data, c, lr, input.layer, output.layer)
            weight.mat <- sim_results$wm
            trainVec <- c(unique(train_data$x))

            # Test on the testing data
            predictions_test <- sapply(test_data$x, function(input) {
                #exam.response(input, c, trainVec, input.layer, output.layer, weight.mat)
                exam.response2(input, c, train_data, input.layer, output.layer, weight.mat)

            })

            # Test on the latter half of the training data
            predictions_train <- sapply(train_data_half$x, function(input) {
                alm.response(input, c, input.layer, output.layer, weight.mat)$mean.response
            })

            # Compute mean squared error for test and latter half of training data
            mse_test <- mean((predictions_test - test_data$y)^2)
            mse_train <- mean((predictions_train - train_data_half$y)^2)

            # Use the average of the two MSEs as the final error
            final_error <- mean(c(mse_test, mse_train), na.rm = TRUE)

            # Update best parameters if current error is lower
            if (!is.na(final_error) && final_error < best_error) {
                best_error <- final_error
                best_params$c <- c
                best_params$lr <- lr
            }
        }
    }
     (test_data <- cbind(test_data, pred = predictions_test))
    print(best_params)
    print(best_error)

    return(best_params)
}

# Example usage
c_values <- seq(0.000001, 0.01, length.out=40)
lr_values <- seq(0.01, 1.5, length.out=20)
best_paramsV <- optimize_params(dsAvg |> filter(condit=="Varied"), c_values, lr_values, input.layer, output.layer)
c = best_paramsV$c; lr = best_paramsV$lr
viewPreds(dsAvg |> filter(condit=="Varied"), c_values, lr_values, input.layer, output.layer)

best_paramsC <- optimize_params(dsAvg |> filter(condit=="Constant"), c_values, lr_values, input.layer, output.layer)
c = best_paramsC$c; lr = best_paramsC$lr
viewPreds(dsAvg |> filter(condit=="Constant"), c_values, lr_values, input.layer, output.layer)

@@

## ALM test
```{r}

dsAvg <- ds |> group_by(condit,expMode2,tr, x) |> summarise(y=mean(y),.groups="keep") 
#dsAvg <- dsAvg |> filter(condit=="Varied")

optimize_params <- function(dsAvg, c_values, lr_values, input.layer, output.layer){
    best_error <- Inf
    best_params <- list(c = NA, lr = NA)

    # Split the data into training and testing sets
    train_data <- dsAvg[dsAvg$expMode2 == "Train", ]
    test_data <- dsAvg[dsAvg$expMode2 == "Test", ] |> group_by(x) |> summarise(y = mean(y))

    # Determine the second half of the training data
    train_data_half <- train_data[(round(nrow(train_data)/2) + 1):nrow(train_data), ]

    # Grid search over c and lr values
    for (c in c_values) {
        for (lr in lr_values) {
            # Train ALM with current parameters
            sim_results <- alm.sim(train_data, c, lr, input.layer, output.layer)
            weight.mat <- sim_results$wm
            trainVec <- c(unique(train_data$x))

            # predictions_train <- sim_results$d[(round(nrow(sim_results$d)/2) + 1):nrow(sim_results$d), ] |> mutate(error=(y-almResp))

             predictions_train <- sim_results$d[2:nrow(sim_results$d),] |> mutate(error=(y-almResp))

            # Test on the testing data
            predictions_test <- sapply(test_data$x, function(input) {
                #exam.response(input, c, trainVec, input.layer, output.layer, weight.mat)
                alm.response(input, c, input.layer, output.layer, weight.mat)$mean.response

            })

            # Compute mean squared error for test and latter half of training data
            mse_test <- mean((predictions_test - test_data$y)^2, na.rm=TRUE)
            mse_train <- mean(predictions_train$error^2)

            # Use the average of the two MSEs as the final error
            final_error <- mean(c( mse_train), na.rm = TRUE)

            # Update best parameters if current error is lower
            if (!is.na(final_error) && final_error < best_error) {
                best_error <- final_error
                best_params$c <- c
                best_params$lr <- lr
            }
        }
    }
     (test_data <- cbind(test_data, pred = predictions_test))
    print(best_params)
    print(best_error)

    return(best_params)
}

# Example usage
c_values <- seq(0.000001, 0.01, length.out=50)
lr_values <- seq(0.01, 1.5, length.out=30)
best_paramsV <- optimize_params(dsAvg |> filter(condit=="Varied"), c_values, lr_values, input.layer, output.layer)
c = best_paramsV$c; lr = best_paramsV$lr
viewPreds(dsAvg |> filter(condit=="Varied"), c, lr, input.layer, output.layer)

best_paramsC <- optimize_params(dsAvg |> filter(condit=="Constant"), c_values, lr_values, input.layer, output.layer)
c = best_paramsC$c; lr = best_paramsC$lr
viewPreds(dsAvg |> filter(condit=="Constant"), c, lr, input.layer, output.layer)

@@

best_paramsV <- optimize_params(dsAvg |> filter(condit=="Varied"), c_values, lr_values, input.layer, output.layer)
$c
[1] 0.001026538

$lr
[1] 0.2452632

[1] 9081.058
> c = best_paramsV$c; lr = best_paramsV$lr
> viewPreds(dsAvg |> filter(condit=="Varied"), c, lr, input.layer, output.layer)
     x         y     pred      pred
1  100  663.0304  329.644  978.7495
2  350  764.2023  561.467  978.7495
3  600  883.9191  793.291  978.7495
4  800 1083.0380  978.749  978.7495
5 1000 1195.5169 1090.704 1090.7037
6 1200 1283.1824 1054.657 1054.6570
> best_paramsC <- optimize_params(dsAvg |> filter(condit=="Constant"), c_values, lr_values, input.layer, output.layer)
c = best_paramsC$c; lr = best_paramsC$lr
$c
[1] 0.0002573846

$lr
[1] 0.01

[1] 1911.687
> c = best_paramsC$c; lr = best_paramsC$lr
> viewPreds(dsAvg |> filter(condit=="Constant"), c, lr, input.layer, output.layer)
     x         y     pred     pred
1  100  526.6764  625.899 920.9065
2  350  666.3083  731.259 920.9065
3  600  779.6170  836.619 920.9065
4  800  980.0433  920.907 920.9065
5 1000 1162.7106 1005.194 920.9065
6 1200 1276.5916 1089.482 920.9065

# ALM-only model prediction
predict_alm_only <- function(input, c, trainData,input.layer, output.layer, weight.mat) {
  alm.response(input, c, input.layer, output.layer, weight.mat)$mean.response
}

# ALM-EXAM model prediction
predict_alm_exam <- function(input, c, trainData, input.layer, output.layer, weight.mat) {
  exam.response2(input, c, trainData, input.layer, output.layer, weight.mat)
}

# ALM-EXAM hybrid model prediction
predict_alm_exam_hybrid <- function(input, c, trainData, input.layer, output.layer, weight.mat) {
  alm_pred <- predict_alm_only(input, c,trainData, input.layer, output.layer, weight.mat)
  exam_pred <- predict_alm_exam(input, c, trainData, input.layer, output.layer, weight.mat)

  # Average the predictions from ALM and EXAM
  (alm_pred + exam_pred) / 2
}

optimize_params <- function(dsAvg, c_values, lr_values, input.layer, output.layer, predict_function) {
    best_error <- Inf
    best_params <- list(c = NA, lr = NA)

    # Split the data into training and testing sets
    train_data <- dsAvg[dsAvg$expMode2 == "Train", ]
    test_data <- dsAvg[dsAvg$expMode2 == "Test", ] |> group_by(x) |> summarise(y=mean(y))

    # Grid search over c and lr values
    for (c in c_values) {
        for (lr in lr_values) {
            # Train ALM with current parameters
            sim_results <- alm.sim(train_data, c, lr, input.layer, output.layer)
            weight.mat <- sim_results$wm

            # Test on the testing data using the provided prediction function
            predictions <- sapply(test_data$x, function(input) {
                predict_function(input, c, train_data, input.layer, output.layer, weight.mat)
            })

            # Compute mean squared error
            mse <- mean((predictions - test_data$y)^2)

            # Update best parameters if current error is lower
            if (!is.na(mse) && mse < best_error) {
                best_error <- mse
                best_params$c <- c
                best_params$lr <- lr
            }
        }
    }

    list(best_params=best_params, best_error=best_error)
}

c_values <- seq(0.000001, 0.01, length.out=50)
lr_values <- seq(0.01, 1.5, length.out=50)

result_alm_only <- optimize_params(dsAvg |> filter(condit=="Varied"), c_values, lr_values, input.layer, output.layer, predict_alm_only)

result_alm_exam <- optimize_params(dsAvg |> filter(condit=="Varied"), c_values, lr_values, input.layer, output.layer, predict_alm_exam)

result_alm_exam_hybrid <- optimize_params(dsAvg |> filter(condit=="Varied"), c_values, lr_values, input.layer, output.layer, predict_alm_exam_hybrid)

map(list(result_alm_only, result_alm_exam, result_alm_exam_hybrid), 
    ~ viewPreds(dsAvg |> filter(condit=="Varied"), .x$best_params$c, .x$best_params$lr, input.layer,output.layer))

@@

[[1]]
     x         y exam2.pred alm.pred
1  100  663.0304    145.955  795.061
2  350  764.2023    377.779  795.061
3  600  883.9191    609.602  795.061
4  800 1083.0380    795.061  795.061
5 1000 1195.5169   1199.920 1199.920
6 1200 1283.1824   1151.507 1151.507

[[2]]
     x         y exam2.pred alm.pred
1  100  663.0304    381.812      NaN
2  350  764.2023    613.635      NaN
3  600  883.9191    845.458 1030.917
4  800 1083.0380   1030.917 1030.917
5 1000 1195.5169   1097.012 1097.012
6 1200 1283.1824   1136.143 1136.143

[[3]]
     x         y exam2.pred alm.pred
1  100  663.0304    356.566 1005.670
2  350  764.2023    588.390 1005.670
3  600  883.9191    820.213 1005.670
4  800 1083.0380   1005.672 1005.672
5 1000 1195.5169   1088.837 1088.837
6 1200 1283.1824   1138.575 1138.575

```{r}
predict_alm_exam_weighted_hybrid <- function(input, c, weight_exam, trainData, input.layer, output.layer, weight.mat) {

  alm_pred <- predict_alm_only(input, c, trainData, input.layer, output.layer, weight.mat)
    exam_pred <- predict_alm_exam(input, c, trainData, input.layer, output.layer, weight.mat)

  # Weighted combination of the predictions from ALM and EXAM
  (1 - weight_exam) * alm_pred + weight_exam * exam_pred
}

optimize_params_weighted <- function(dsAvg, c_values, lr_values, weight_exam_values, input.layer, output.layer) {
    best_error <- Inf
    best_params <- list(c = NA, lr = NA, weight_exam = NA)
    error_data <- data.frame() # Empty data frame to store errors

    # Split the data into training and testing sets
    train_data <- dsAvg[dsAvg$expMode2 == "Train", ]
    test_data <- dsAvg[dsAvg$expMode2 == "Test", ] |> group_by(x) |> summarise(y=mean(y))

    # Grid search over c, lr, and weight_exam values
    for (c in c_values) {
        for (lr in lr_values) {
            for (weight_exam in weight_exam_values) {
                # Train ALM with current parameters
                sim_results <- alm.sim(train_data, c, lr, input.layer, output.layer)
                weight.mat <- sim_results$wm

                # Test on the testing data using the weighted hybrid prediction function
                predictions <- sapply(test_data$x, function(input) {
                    predict_alm_exam_weighted_hybrid(input, c, weight_exam, train_data, input.layer, output.layer, weight.mat)
                })

                # Compute mean squared error
                mse <- mean((predictions - test_data$y)^2)
                error_data <- rbind(error_data, data.frame(c=c, lr=lr, weight_exam=weight_exam, mse=mse))

                # Update best parameters if current error is lower
                if (!is.na(mse) && mse < best_error) {
                    best_error <- mse
                    best_params$c <- c
                    best_params$lr <- lr
                    best_params$weight_exam <- weight_exam
                }
            }
        }
    }

    list(best_params=best_params, best_error=best_error, error_data=error_data)
}

# Example usage
weight_exam_values <- seq(.2, .8, length.out=6) # or any other desired sequence
c_values <- seq(0.0001, 0.01, length.out=40)
lr_values <- seq(0.01, 1.5, length.out=30)
result_weighted_hybrid <- optimize_params_weighted(dsAvg |> filter(condit=="Varied"), c_values, lr_values, weight_exam_values, input.layer, output.layer)
c = result_weighted_hybrid$best_params$c; lr = result_weighted_hybrid$best_params$lr; weight_exam = result_weighted_hybrid$best_params$weight_exam;
viewPreds(dsAvg |> filter(condit=="Varied"),c, lr, input.layer, output.layer, weight_exam)

error_data <- result_weighted_hybrid$error_data

# Visualize the error surface
p <- ggplot(error_data, aes(x=c, y=lr, z=mse)) +
    geom_line(aes(col=mse)) +
    scale_fill_gradient(low="blue", high="red") +
    theme_void() +
    # use viridis color palette
    scale_color_viridis_c(option="plasma") +
    labs(x="c", y="lr", z="mse") +
    guides(color=FALSE) +
    theme(legend.position="none")

plot_ly(error_data, x = ~c, y = ~lr, z = ~mse, type = "scatter3d", mode = "markers", color = ~mse, colors = c("blue", "red")) %>%
  layout(scene = list(xaxis = list(title = "c"),
                      yaxis = list(title = "lr"),
                      zaxis = list(title = "mse")))

$best_params
$best_params$c
[1] 0.0004091224

$best_params$lr
[1] 0.03855897

$best_params$weight_exam
[1] 0.5555556

$best_error
[1] 6442.535

     x         y exam2.pred alm.pred hybrid.pred  alm.pred hybridFit.pred
1  100  663.0304    359.061 1008.143    683.6021  683.6021       647.5420
2  350  764.2023    590.884 1008.143    799.5136  799.5136       776.3326
3  600  883.9191    822.707 1008.143    915.4251  915.4251       905.1231
4  800 1083.0380   1008.166 1008.166   1008.1660 1008.1660      1008.1660
5 1000 1195.5169   1089.756 1089.756   1089.7562 1089.7562      1089.7562
6 1200 1283.1824   1138.174 1138.174   1138.1740 1138.1740      1138.1740

```{r}
weight_exam_values <- seq(0, 1, length.out=10) # or any other desired sequence
c_values <- seq(0.00001, 0.01, length.out=50)
lr_values <- seq(0.00001, 1.5, length.out=40)
result_weighted_hybridC <- optimize_params_weighted(dsAvg |> filter(condit=="Constant"), c_values, lr_values, weight_exam_values, input.layer, output.layer)

best_params <- result_weighted_hybridC$best_params

c = best_params$c; lr = best_params$lr; weight_exam = best_params$weight_exam;
result_weighted_hybridC$best_error
viewPreds(dsAvg |> filter(condit=="Constant"),c, lr, input.layer, output.layer, weight_exam)

result_weighted_hybridC$best_error
[1] 13427.37
> best_params
$c
[1] 0.0002138776

$lr
[1] 1e-05

$weight_exam
[1] 1

> viewPreds(dsAvg |> filter(condit=="Constant"),c, lr, input.layer, output.layer, weight_exam)
     x         y exam2.pred alm.pred hybrid.pred  alm.pred hybridFit.pred
1  100  526.6764    626.082 921.0898    773.5859  773.5859        626.082
2  350  666.3083    731.442 921.0898    826.2659  826.2659        731.442
3  600  779.6170    836.802 921.0898    878.9459  878.9459        836.802
4  800  980.0433    921.090 921.0898    921.0899  921.0899        921.090
5 1000 1162.7106   1005.378 921.0898    963.2339  963.2339       1005.378
6 1200 1276.5916   1089.666 921.0898   1005.3779 1005.3779       1089.666

@@

weight_exam_values <- seq(0, 1, length.out=15) # or any other desired sequence
c_values <- seq(0.000001, 0.5, length.out=50)
lr_values <- seq(0.000001, 1.5, length.out=50)

all_results_weighted_hybrid <- optimize_params_weighted_individual(ds, c_values, lr_values, weight_exam_values, input.layer, output.layer)

saveRDS(all_results_weighted_hybrid,file=paste0(here::here("data/model_cache/indv_hybrid_fits2.rds")))

@@

viewPreds <- function(dsAvg, c, lr, input.layer, output.layer,weightExam=.5)
{
    train_data <- dsAvg[dsAvg$expMode2 == "Train", ]
    test_data <- dsAvg[dsAvg$expMode2 == "Test", ] |> group_by(x) |> summarise(y = mean(y))
     sim_results <- alm.sim(train_data, c, lr, input.layer, output.layer)
      weight.mat <- sim_results$wm
      trainVec <- c(unique(train_data$x))

    # predictions_test <- sapply(test_data$x, function(input) {
    #       exam.response(input, c, trainVec, input.layer, output.layer, weight.mat)
    #         })
    # (test_data <- cbind(test_data, exam.pred = predictions_test))  
    # 
    #  trainVec <- c(100,unique(train_data$x))
    # predictions_test <- sapply(test_data$x, function(input) {
    #       exam.response(input, c, trainVec, input.layer, output.layer, weight.mat)
    #         })
    # (test_data <- cbind(test_data, exam.pred0 = predictions_test))

      predictions_test <- sapply(test_data$x, function(input) {
          exam.response2(input, c, train_data, input.layer, output.layer, weight.mat)
            })
    (test_data <- cbind(test_data, exam2.pred = predictions_test))

      predictions_test <- sapply(test_data$x, function(input) {
            #exam.response(input, c, trainVec, input.layer, output.layer, weight.mat)
            alm.response(input, c, input.layer, output.layer, weight.mat)$mean.response
            })

    (test_data <- cbind(test_data, alm.pred = predictions_test))
  predictions_test <- sapply(test_data$x, function(input) {
            #exam.response(input, c, trainVec, input.layer, output.layer, weight.mat)
             predict_alm_exam_hybrid(input, c, train_data,input.layer, output.layer, weight.mat)
            })
    (test_data <- cbind(test_data, hybrid.pred = predictions_test))

    (test_data <- cbind(test_data, alm.pred = predictions_test))
  predictions_test <- sapply(test_data$x, function(input) {
            #exam.response(input, c, trainVec, input.layer, output.layer, weight.mat)
             predict_alm_exam_weighted_hybrid(input, c, weightExam, train_data,input.layer, output.layer, weight.mat)
            })
    (test_data <- cbind(test_data, hybridFit.pred = predictions_test))

}

```{r}
i=1
dat=train_data
input=800; corRep=975
c=.001; lr=.5
 weight.mat <- matrix(0.01, nrow = length(output.layer), ncol = length(input.layer))
  xt <- dat$x
  n <- nrow(dat)
  st <- numeric(n) # Initialize the vector to store mean responses
  for(i in 1:n) {
    trial <- alm.trial(dat$x[i], dat$y[i], c, lr, input.layer, output.layer, weight.mat)
    weight.mat <- trial$weight.mat
    st[i] <- trial$mean.response
  }
  dat <- cbind(dat,almResp=st)

alm_resp <- alm.response(input, c, input.layer,output.layer, weight.mat)

input.activation <- exp(-c * (input.layer - input)^2) / sum(exp(-c * (input.layer - input)^2))
output.activation <- weight.mat %*% input.activation
output.probability <- output.activation / sum(output.activation)
mean.response <- sum(output.layer * output.probability)
```

>>>>

```{r}
# load and view data
pacman::p_load(tidyverse,lme4,future,furrr,patchwork,here, furrr, future, pander)
purrr::walk(here(c("Functions/Display_Functions.R", "Functions/alm_core.R","Functions/misc_model_funs.R")),source)
select <- dplyr::select; mutate <- dplyr::mutate

ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))

dsAvg <- ds |> group_by(condit,expMode2,tr, x) |> 
  summarise(y=mean(y),.groups="keep")

vAvg <- dsAvg |> filter(condit=="Varied")
cAvg <- dsAvg |> filter(condit=="Constant")

@@

i1 <- ds |> filter(id=="3")

default.layer <- c(100,350,600,800,1000,1200)

output.layer <- adjust_layer(default.layer,k=3)

@@

a_testOnly=list(pred_dat="test_avg",pred_fun="alm.responseOnly",loss_fun="RMSE",loss_data="test_error")
a_trainOnly=list(pred_dat="test_avg",pred_fun="alm.responseOnly",loss_fun="RMSE",loss_data="train_error")
a_testTrain=list(pred_dat="test_avg",pred_fun="alm.responseOnly",loss_fun="RMSE",loss_data="test_error+train_error")

e_testOnly=list(pred_dat="test_avg",pred_fun="exam.response",loss_fun="RMSE",loss_data="test_error")
e_trainOnly=list(pred_dat="test_avg",pred_fun="exam.response",loss_fun="RMSE",loss_data="train_error")
e_testTrain=list(pred_dat="test_avg",pred_fun="exam.response",loss_fun="RMSE",loss_data="test_error+train_error")

h_testOnly=list(pred_dat="test_avg",pred_fun="predict_alm_exam_weighted_hybrid",loss_fun="RMSE",loss_data="test_error")
h_trainOnly=list(pred_dat="test_avg",pred_fun="predict_alm_exam_weighted_hybrid",loss_fun="RMSE",loss_data="train_error")
h_testTrain=list(pred_dat="test_avg",pred_fun="predict_alm_exam_weighted_hybrid",loss_fun="RMSE",loss_data="test_error+train_error")

hybrid_te_v <- wrap_grid(vAvg, c_values, lr_values, input.layer, output.layer,predParams=e_testOnly)
hybrid_tetr_v <- wrap_grid(vAvg, c_values, lr_values, input.layer, output.layer,predParams=e_testTrain)
hybrid_tr_v<- wrap_grid(vAvg, c_values, lr_values, input.layer, output.layer,predParams=e_trainOnly)

```{r}
plan(multisession)

c_values <- seq(0.000001, 1.0, length.out=120)
lr_values <- seq(0.0000001, 4.0, length.out=200)

list2env(readRDS(here::here("data/model_cache/con_group_exam_fits.rds")), envir = .GlobalEnv)
list2env(readRDS(here::here("data/model_cache/var_group_exam_fits.rds")), envir = .GlobalEnv)

@@

@@

## Varied Testing
```{r fig.height=11, fig.width=11}

vte <- pluck(a_te_v, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex_te_v, "test") |> pull(pred)) |>  
  pivot_longer(Observed:EXAM, names_to="Model", values_to = "vx") |> 
  ggplot(aes(x,vx,fill=Model,col=Model,shape=Model)) +geom_point() + 
  scale_x_continuous(breaks=sort(unique(ds$x)), labels=sort(unique(ds$x)))+ylim(0,1500) +
  theme(legend.title = element_blank(), legend.position="top") +ggtitle("Optimize for Test Only")

vtetr <- pluck(a_tetr_v, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex_tetr_v, "test") |> pull(pred)) |>  
  pivot_longer(Observed:EXAM, names_to="Model", values_to = "vx") |> 
  ggplot(aes(x,vx,fill=Model,col=Model,shape=Model)) +geom_point() + 
  scale_x_continuous(breaks=sort(unique(ds$x)), labels=sort(unique(ds$x)))+ylim(0,1500) +
  theme(legend.title = element_blank(), legend.position="top") +ggtitle("Optimize for Test and Train")

vtr <- pluck(a_tr_v, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex_tr_v, "test") |> pull(pred)) |>  
  pivot_longer(Observed:EXAM, names_to="Model", values_to = "vx") |> 
  ggplot(aes(x,vx,fill=Model,col=Model,shape=Model)) +geom_point(alpha=.5) + 
  scale_x_continuous(breaks=sort(unique(ds$x)), labels=sort(unique(ds$x)))+ ylim(0,1500) +
  theme(legend.title = element_blank(), legend.position="top") +ggtitle("Optimize for Train Only")

vte/vtetr/vtr

tvtr<- pluck(a_tr_v, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex_tr_v, "test") |> pull(pred))

tvtetr<- pluck(a_tetr_v, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex_tetr_v, "test") |> pull(pred))

tvte<- pluck(a_te_v, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex_te_v, "test") |> pull(pred))

pander("Varied-fit to test only") + pander(tvte) +
  pander("Varied - fit to train and test")+ pander(tvtetr)+
  pander("Varired-fit to train only")+ pander(tvtr)
```

## Constant Testing
```{r fig.height=11, fig.width=11}

cte <- pluck(a_te_c, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex0_te_c, "test") |> pull(pred)) |>  
  pivot_longer(Observed:EXAM, names_to="Model", values_to = "vx") |> 
  ggplot(aes(x,vx,fill=Model,col=Model,shape=Model)) +geom_point() + 
  scale_x_continuous(breaks=sort(unique(ds$x)), labels=sort(unique(ds$x)))+ylim(0,1500) +
  theme(legend.title = element_blank(), legend.position="top") +ggtitle("Optimize for Test Only")

ctetr <- pluck(a_tetr_c, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex0_tetr_c, "test") |> pull(pred)) |>  
  pivot_longer(Observed:EXAM, names_to="Model", values_to = "vx") |> 
  ggplot(aes(x,vx,fill=Model,col=Model,shape=Model)) +geom_point() + 
  scale_x_continuous(breaks=sort(unique(ds$x)), labels=sort(unique(ds$x)))+ylim(0,1500) +
  theme(legend.title = element_blank(), legend.position="top") +ggtitle("Optimize for Test and Train")

ctr <- pluck(a_tr_c, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex0_tr_c, "test") |> pull(pred)) |>  
  pivot_longer(Observed:EXAM, names_to="Model", values_to = "vx") |> 
  ggplot(aes(x,vx,fill=Model,col=Model,shape=Model)) +geom_point() + 
  scale_x_continuous(breaks=sort(unique(ds$x)), labels=sort(unique(ds$x)))+ ylim(0,1500) +
  theme(legend.title = element_blank(), legend.position="top") +ggtitle("Optimize for Train Only")

cte/ctetr/ctr

tctr<- pluck(a_tr_c, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex0_tr_c, "test") |> pull(pred))

tctetr<- pluck(a_tetr_c, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex0_tetr_c, "test") |> pull(pred))

tcte<- pluck(a_te_c, "test") |> rename(ALM=pred,Observed=y) %>% 
  cbind(.,EXAM=pluck(ex0_te_c, "test") |> pull(pred))

pander("Constant-fit to test only") + pander(tcte) +
  pander("Constant - fit to train and test")+ pander(tctetr)+
  pander("Constant-fit to train only")+ pander(tctr)

@@

#ALM learning curves
```{r}
pluck(a_te_v, "train") |> pivot_longer(y:almResp, names_to="Resp", values_to = "vx") |> 
  mutate(dev=x-vx,abs_dev=abs(x-vx)) |>
  learn_curve_plot(tr, vx, Resp,facet_var=x, groupVec=Resp,nbins=8)

pluck(a_te_v, "train") |> pivot_longer(y:almResp, names_to="Resp", values_to = "vx") |> 
  mutate(dev=x-vx,abs_dev=abs(x-vx)) |>
  ungroup() %>%
  gather(key = "variable", value = "y_value", dev, abs_dev, vx) %>%
  group_by(variable) %>%
  group_map(~ learn_curve_plot(.x, x_var = tr, y_var = y_value, color_var = Resp, facet_var = x, groupVec = Resp, nbins = 8, y_label = .y$variable), .keep = TRUE)

list(a_tr_v, a_te_v,a_tetr_v) |> map( ~{pluck(.x, "train") |> pivot_longer(y:almResp, names_to="Resp", values_to = "vx") |> mutate(dev=x-vx,abs_dev=abs(x-vx)) |>
  ungroup() %>%
  gather(key = "variable", value = "y_value", dev, abs_dev, vx) %>%
  group_by(variable) %>%
  group_map(~ learn_curve_plot(.x, x_var = tr, y_var = y_value, color_var = Resp, facet_var = x, groupVec = Resp, nbins = 8, y_label = .y$variable), .keep = TRUE) })

@@

# EXAM fit learning curves
```{r}

pluck(ex_te_v, "train") |> pivot_longer(y:almResp, names_to="Resp", values_to = "vx") |> 
  mutate(dev=x-vx,abs_dev=abs(x-vx)) |>
  learn_curve_plot(tr, vx, Resp,facet_var=x, groupVec=Resp,nbins=8)

pluck(ex_te_v, "train") |> pivot_longer(y:almResp, names_to="Resp", values_to = "vx") |> 
  mutate(dev=x-vx,abs_dev=abs(x-vx)) |>
  ungroup() %>%
  gather(key = "variable", value = "y_value", dev, abs_dev, vx) %>%
  group_by(variable) %>%
  group_map(~ learn_curve_plot(.x, x_var = tr, y_var = y_value, color_var = Resp, facet_var = x, groupVec = Resp, nbins = 8, y_label = .y$variable), .keep = TRUE)

@@

@@

@@
